---
title: "Feather River Mini Snorkel Data QC"
author: "Erin Cain"
date: "03/24"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.width=15, fig.height=10)
library(tidyverse)
library(lubridate)
library(googleCloudStorageR)
library(ggplot2)
library(scales)

colors_small <-  c("#9A8822", "#F5CDB4", "#F8AFA8", "#FDDDA0", "#74A089", #Royal 2
                   "#899DA4", "#C93312", "#DC863B" # royal 1 (- 3)
)
```

# Feather River Mini Snorkel Data

## Description of Monitoring Data

**Timeframe:** March 2001 - August 2001

**Completeness of Record throughout timeframe:** fairly complete 

**Sampling Location:** Feather River 

**Data Contact:** Ryon Kurth 

## Questions 

- what is distance 
- what is t_code

## Source Database pull

```{r, echo=TRUE}
source("data-raw/query_4mac.R")
```

Read in data from google cloud, glimpse raw data and domain description sheet: 
```{r}
# read in data to clean 
all_fish_obs |> glimpse()
```

## Data transformations

### first table reviewd is the All Fish Observation Table 
All of the substrate and cover lookups are not true look ups. Substrate and cover column indicate a percentage of cover or substrate of each type. I updated the column names to reflect this and utilized the lookup tables to understand which substrate or cover type each column was referring to. 

Columns removed: 
- SpecAge removed - just a combination of species and age

```{r}
# For different excel sheets for each year read in and combine years here
joined_fish_obs <- all_fish_obs |> 
  left_join(species_code_lookup) |> 
  select(-c(SpeciesCodeID, SpeciesCode)) |>
  rename(percent_fine_substrate = Sub1, 
         percent_sand_substrate = Sub2, 
         percent_small_gravel_substrate = Sub3, 
         percent_large_gravel_substrate = Sub4, 
         percent_cobble_substrate = Sub5, 
         percent_boulder_substrate = Sub6,
         percent_no_cover_inchannel = IcovA, 
         percent_small_woody_cover_inchannel = IcovB, 
         percent_large_woody_cover_inchannel = IcovC, 
         percent_submerged_aquatic_veg_inchannel = IcovE, 
         percent_undercut_bank = IcovF,
         percent_no_cover_overhead = Ocov0,
         percent_cover_half_meter_overhead = Ocov1, 
         percent_cover_more_than_half_meter_overhead = Ocov2)|> 
  # Clean in separate file
  # left_join(location_table |> mutate(PhysDataTblID = as.numeric(PhysDataTblID),
  #                                    Location = as.character(Location))) |> 
  janitor::clean_names() |> 
  select(-spec_age) |> #same as species + age
  left_join(cgu_code_lookup |> mutate(CGUCode = tolower(CGUCode)), by = c("cgu" = "CGUCode")) |> 
  select(-cgu, -CGUCodeID) |> 
  rename(channel_geomorphic_unit = CGU) |> 
  glimpse()
  
```

## Explore Numeric Variables: {.tabset}

```{r}
# Filter clean data to show only numeric variables 
joined_fish_obs %>% select_if(is.numeric) %>% colnames()

```
### Variable: `count`

**Plotting Count over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
joined_fish_obs %>% 
  ggplot(aes(x = date, y = count, group = 1))+
  # geom_line()+
  geom_point(aes(x=date, y = count))+
  theme_minimal()+
  labs(title = "Count over Time",
       y = "Number of Fish Observations")
```

**Numeric Summary of Count over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$count)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$count) |> sum()` NA values


### Variable: `date`

**Plotting Date over Period of Record**

All observations are from 2001 

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
joined_fish_obs %>% 
  ggplot(aes(x = date, y = TRUE, group = 1))+
  # geom_line()+
  geom_point(aes(x=date, y = TRUE))+
  theme_minimal()+
  labs(title = "Dates Surveyed",
       y = "")
```

**Numeric Summary of Count over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$date)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$date) |> sum()` NA values


### Variable: `river_mile``

**Plotting river_mile over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
joined_fish_obs %>% 
  ggplot(aes(x = river_mile, y = date, group = 1))+
  # geom_line()+
  geom_point(aes(x=river_mile, y = date))+
  theme_minimal()+
  labs(title = "Dates that River miles Surveyed",
       y = "")
```

**Numeric Summary of river_mile over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$river_mile)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$river_mile) |> sum()` NA values

### Variable: `fl_mm`

**Plotting fl_mm over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
joined_fish_obs |> 
  # filter(fork_length < 250) %>% # filter out 13 points so we can more clearly see distribution
  ggplot(aes(x = fl_mm)) + 
  geom_histogram(breaks=seq(0, 200, by=2)) + 
  scale_x_continuous(breaks=seq(0, 200, by=25)) +
  theme_minimal() +
  labs(title = "Fork length distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```
```{r}
joined_fish_obs %>% 
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = fl_mm, y = species)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Fork length summarized by species") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

```


**Numeric Summary of fl_mm over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$fl_mm)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$fl_mm) |> sum()` NA values


### Variable: `dist_to_bottom`

**Plotting dist_to_bottom distribution**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
joined_fish_obs |> 
  # filter(fork_length < 250) %>% # filter out 13 points so we can more clearly see distribution
  ggplot(aes(x = dist_to_bottom)) + 
  geom_histogram(breaks=seq(0, 200, by=2)) + 
  scale_x_continuous(breaks=seq(0, 200, by=25)) +
  theme_minimal() +
  labs(title = "Distance to Bottom") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

```{r}

joined_fish_obs %>% 
  mutate(year = as.factor(year(date))) %>%
  ggplot(aes(x = dist_to_bottom, y = as.factor(river_mile))) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Distance to bottom summarized by river_mile") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of dist_to_bottom over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$river_mile)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$river_mile) |> sum()` NA values


### Variable: `focal_velocity` & `velocity`

**Plotting velocities over Period of Record**

```{r}
joined_fish_obs |> 
  # filter(fork_length < 250) %>% # filter out 13 points so we can more clearly see distribution
  ggplot(aes(x = focal_velocity)) + 
  geom_histogram(breaks=seq(0, 10, by=1), group = "Focal velocity", fill = "blue", alpha = .2 ) + 
  geom_histogram(aes(x = velocity), breaks=seq(0, 10, by=1), group = "Focal velocity", fill = "red", alpha = .2 ) + 
  scale_x_continuous(breaks=seq(0, 10, by=1)) +
  theme_minimal() +
  labs(title = "Focal Velocity vs ") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

Looks like velocity and focal velocity have similar distributions

**Numeric Summary of focal_velocity over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$focal_velocity)
summary(joined_fish_obs$velocity)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$focal_velocity) |> sum()` NA values
There are `r is.na(joined_fish_obs$velocity) |> sum()` NA values

### Variable: `t_code`

**Plotting t_code over Period of Record**

```{r}
joined_fish_obs |> 
  # filter(fork_length < 250) %>% # filter out 13 points so we can more clearly see distribution
  ggplot(aes(x = t_code)) + 
  geom_histogram(breaks=seq(0, 30, by=1)) + 
  scale_x_continuous(breaks=seq(0, 30, by=1)) +
  theme_minimal() +
  labs(title = "T code distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of t_code over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$t_code)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$t_code) |> sum()` NA values

### Variable: `depth`

**Plotting depth over Period of Record**

```{r}
joined_fish_obs |> 
  # filter(fork_length < 250) %>% # filter out 13 points so we can more clearly see distribution
  ggplot(aes(x = depth)) + 
  geom_histogram(breaks=seq(0, 250, by=1)) + 
  scale_x_continuous(breaks=seq(0, 200, by=50)) +
  theme_minimal() +
  labs(title = "T code distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of depth over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$depth)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$t_code) |> sum()` NA values

### Variable: `substrate `

**Plotting substrate over Period of Record**

```{r}
# Make whatever plot is appropriate 
# maybe 2+ plots are appropriate
joined_fish_obs |> 
  arrange(river_mile) |> 
  mutate(obs_id = 1:nrow(joined_fish_obs)) |> 
  select(obs_id, 
         percent_fine_substrate,
         percent_sand_substrate,
         percent_small_gravel_substrate,
         percent_large_gravel_substrate,
         percent_cobble_substrate,
         percent_boulder_substrate) |> 
  pivot_longer(cols = percent_fine_substrate:percent_boulder_substrate, 
               names_to = "substrate_type", values_to = "percent") |> 
  ggplot(aes(x = obs_id, y = percent, fill = substrate_type)) +
  geom_col() +
  scale_fill_manual(values = colors_small) +
  theme_minimal() +
  theme(legend.position = "bottom")


```

**Numeric Summary of substrate over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$percent_fine_substrate)
summary(joined_fish_obs$percent_sand_substrate)
summary(joined_fish_obs$percent_small_gravel_substrate)
summary(joined_fish_obs$percent_large_gravel_substrate)
summary(joined_fish_obs$percent_boulder_substrate)
summary(joined_fish_obs$percent_cobble_substrate)

```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$percent_fine_substrate) |> sum()` NA values
There are `r is.na(joined_fish_obs$percent_sand_substrate) |> sum()` NA values
There are `r is.na(joined_fish_obs$percent_small_gravel_substrate) |> sum()` NA values
There are `r is.na(joined_fish_obs$percent_small_gravel_substrate) |> sum()` NA values
There are `r is.na(joined_fish_obs$percent_boulder_substrate) |> sum()` NA values
There are `r is.na(joined_fish_obs$percent_cobble_substrate) |> sum()` NA values

### Variable: `inchannel cover`

**Plotting inchannel cover over Period of Record**

Notes: 
- some cover totals less than 100%

```{r}
joined_fish_obs |> 
  arrange(river_mile) |> 
  mutate(obs_id = 1:nrow(joined_fish_obs)) |> 
  select(obs_id, 
         percent_no_cover_inchannel,
         percent_small_woody_cover_inchannel,
         percent_large_woody_cover_inchannel,
         percent_submerged_aquatic_veg_inchannel) |> 
  pivot_longer(cols = percent_no_cover_inchannel:percent_submerged_aquatic_veg_inchannel, 
               names_to = "inchannel_cover_type", values_to = "percent") |> 
  ggplot(aes(x = obs_id, y = percent, fill = inchannel_cover_type)) +
  geom_col() +
  scale_fill_manual(values = colors_small[4:7]) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

**Numeric Summary of inchannel cover over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$percent_no_cover_inchannel)
summary(joined_fish_obs$percent_small_woody_cover_inchannel)
summary(joined_fish_obs$percent_large_woody_cover_inchannel)
summary(joined_fish_obs$percent_submerged_aquatic_veg_inchannel)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$percent_no_cover_inchannel) |> sum()` NA values
There are `r is.na(joined_fish_obs$percent_small_woody_cover_inchannel) |> sum()` NA values
There are `r is.na(joined_fish_obs$percent_large_woody_cover_inchannel) |> sum()` NA values
There are `r is.na(joined_fish_obs$percent_submerged_aquatic_veg_inchannel) |> sum()` NA values

### Variable: `overhead cover`

**Plotting overhead cover over Period of Record**

Notes: 
- some cover totals more than 100%

```{r}
joined_fish_obs |> 
  arrange(river_mile) |> 
  mutate(obs_id = 1:nrow(joined_fish_obs)) |> 
  select(obs_id, 
         percent_undercut_bank,
         percent_no_cover_overhead,
         percent_cover_half_meter_overhead,
         percent_cover_more_than_half_meter_overhead) |> 
  pivot_longer(cols = percent_undercut_bank:percent_cover_more_than_half_meter_overhead, 
               names_to = "overhead_cover_type", values_to = "percent") |> 
  ggplot(aes(x = obs_id, y = percent, fill = overhead_cover_type)) +
  geom_col() +
  scale_fill_manual(values = colors_small[4:7]) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

**Numeric Summary of overhead cover over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$percent_undercut_bank)
summary(joined_fish_obs$percent_no_cover_overhead)
summary(joined_fish_obs$percent_cover_half_meter_overhead)
summary(joined_fish_obs$percent_cover_more_than_half_meter_overhead)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$percent_undercut_bank) |> sum()` NA values
There are `r is.na(joined_fish_obs$percent_no_cover_overhead) |> sum()` NA values
There are `r is.na(joined_fish_obs$percent_cover_half_meter_overhead) |> sum()` NA values
There are `r is.na(joined_fish_obs$percent_cover_more_than_half_meter_overhead) |> sum()` NA values

### Variable: `sur_turb`

**Plotting sur_turb over Period of Record**

```{r}
joined_fish_obs |> 
  # filter(fork_length < 250) %>% # filter out 13 points so we can more clearly see distribution
  ggplot(aes(x = sur_turb)) + 
  geom_histogram(breaks=seq(0, 75, by=1)) + 
  scale_x_continuous(breaks=seq(0, 75, by=5)) +
  theme_minimal() +
  labs(title = "Surface turbidity distribution") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of sur_turb over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$sur_turb)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$sur_turb) |> sum()` NA values

### Variable: `distance`

**Plotting distance over Period of Record**

```{r}
joined_fish_obs |> 
  # filter(fork_length < 250) %>% # filter out 13 points so we can more clearly see distribution
  ggplot(aes(x = distance)) + 
  geom_histogram(breaks=seq(0, 5, by=1)) + 
  scale_x_continuous(breaks=seq(0, 5, by=1)) +
  theme_minimal() +
  labs(title = "Distance") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of distance over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$distance)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$distance) |> sum()` NA values

### Variable: `fish_age`

**Plotting fish_age over Period of Record**

```{r}
joined_fish_obs |> 
  # filter(fork_length < 250) %>% # filter out 13 points so we can more clearly see distribution
  ggplot(aes(x = fish_age)) + 
  geom_histogram(breaks=seq(0, 5, by=.25)) + 
  scale_x_continuous(breaks=seq(0, 5, by=1)) +
  theme_minimal() +
  labs(title = "Distance") + 
  theme(text = element_text(size = 18),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
```

**Numeric Summary of fish_age over Period of Record**

```{r}
# Table with summary statistics
summary(joined_fish_obs$fish_age)
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$fish_age) |> sum()` NA values

## Explore Categorical variables: {.tabset}



```{r}
# Filter clean data to show only categorical variables 
joined_fish_obs %>% select_if(is.character) %>% colnames()
```



### Variable: `location`
```{r}
table(joined_fish_obs$location) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

```{r}
# Fix any inconsistencies with categorical variables
joined_fish_obs$location <- tolower(joined_fish_obs$location)
table(joined_fish_obs$location) 

# FIX aleck riffle and alec riffle
# FIX hatchery  ditch and hatchery ditch
```


**NA and Unknown Values**

There are `r is.na(joined_fish_obs$location) |> sum()` NA values

### Variable: `species`
```{r}
table(joined_fish_obs$species) 
```

Fix inconsistencies with spelling, capitalization, and abbreviations. 

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$species) |> sum()` NA values

### Variable: `channel_geomorphic_unit`
```{r}
table(joined_fish_obs$channel_geomorphic_unit) 
```

**NA and Unknown Values**

There are `r is.na(joined_fish_obs$channel_geomorphic_unit) |> sum()` NA values

## Summary of identified issues

* percent cover that is > or < 100% 
* location variables that I need to change 


## Save cleaned data to data/

```{r}
# Save to data folder
# Name file [watershed]_[data type].csv
```